Optimal Light Curve Attitude Inversion with Measurement Noise: Two Case Studies

### 2025-03-09 paper.tex
added plots, i guess

\usepackage{times}
\usepackage[numbers]{natbib}
\usepackage{graphicx}

\usepackage{bm} % for bold math
\usepackage{mathtools} % for \prescript
\newcommand{\vctr}[1]{\bm{#1}}
\newcommand{\unitv}[1]{\hat{\vctr{#1}}}
\newcommand{\preup}[1]{\prescript{#1}{}{}}
\newcommand{\rf}[1]{\mathcal{\MakeUppercase{#1}}}
\newcommand{\dcm}[1]{\left[\rf{#1}\right]}
\newcommand{\vrf}[2]{\preup{\rf{#1}}\vctr{#2}}
\newcommand{\mbar}[0]{\;\middle|\;}
\newcommand{\prf}[1]{\preup{\rf{#1}}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

\newcommand{\matcp}[1]{\left[#1 \times\right]}
\newcommand{\rthree}[0]{$\mathbb{R}^3$\:}
\newcommand{\sthree}[0]{$\mathbb{S}^3$\:}
\newcommand{\sothree}[0]{$SO_3$}
\newcommand{\pogslla}[0]{$32.900^\circ \textrm{ N}, -105.533^\circ \textrm{ W}$}

@@

% \title{Optimal Full State Light Curve Attitude Inversion: Two Real Case Studies}
\title{Optimal Light Curve Attitude Inversion with Measurement Noise: Two Case Studies}
\author{Liam Robinson}
\author{Carolin Frueh}
\affil{Purdue University, West Lafayette, United States, Email: \texttt{\{robin502, cfrueh\}$@$purdue.edu}}

@@

\section{Inversion Method}

@@

\section{Computing the Light Curve}

@@

\subsection{The Light Curve}

@@

The variance of a space object's light curve whose mean is determined by Equation \ref{eq:general_bright} is a combination of many independent stochastic processes. These distributions have variances $\sigma^2_\text{sensor}$ due to the sensor's integration and readout effects, $\sigma^2_\text{flat}$ from sensor flat-fielding effects, scintillation noise $\bar{S}_{\text{SO},k} \sigma^2_{Y,k}$ \cite{osborn2015}, Poisson signal shot noise $\lambda_{\text{shot},k}$, and Poisson background noise $\lambda_{\text{back},k}$. The sum of these variances yields the total signal variance in ADU:

@@

\section{Real Data Considerations}

\section{Results}

@@

\begin{table}[ht]
  \centering
  \caption{Delta I rocket body observation parameters}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  Telescope location & \pogslla \\ \hline
  Target COSPAR ID & 1984-115C \\ \hline
  First obs.\ (UTC) & Mar 2, 2025 01:53:30.251 \\ \hline
  Light curve duration & $7.5$ minutes \\ \hline
  Observations & $100$ ($92$ extracted) \\ \hline
  Obs.\ frequency & $0.222$ Hz \\ \hline
  Integration time & $3$ seconds \\ \hline
  \end{tabular}
  \label{tb:case1_in}
\end{table}

@@

\begin{table}[ht]
  \centering
  \caption{ECHOSTAR II observation parameters}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  Telescope location & \pogslla \\ \hline
  Target COSPAR ID & 1996-055A \\ \hline
  First obs.\ (UTC) & Feb 26, 2025 04:16:12.279 \\ \hline
  Light curve duration & $19.5$ minutes \\ \hline
  Observations & $456$ ($395$ extracted) \\ \hline
  Obs.\ frequency & $0.388$ Hz \\ \hline
  Integration time & $1$ seconds \\ \hline
  \end{tabular}
  \label{tb:case2_in}
\end{table}

@@

\begin{table}[ht]
  \centering
  \caption{ECHOSTAR II inversion assumptions}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  Aluminum BRDF & $C_d=0.22$, $C_s=0.4$, $n=5$ \\ \hline
  MLI BRDF & $C_d=0.05$, $C_s=0.24$, $n=20$ \\ \hline
  Solar panel BRDF & $C_d=0.05$, $C_s=0.13$, $n=10$ \\ \hline
  $|\omega(0)|_\text{max}$ & $2.32$ $\text{deg} / \text{s}^{-1}$ \\ \hline
  \end{tabular}
  \label{tb:case2_ass}
\end{table}

\section*{Acknowledgments}

@@

\clearpage
\bibliography{cache/refs.bib}
\end{document}

>>>>

\usepackage{times}
\usepackage[numbers]{natbib}
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{bm} % for bold math
\usepackage{mathtools} % for \prescript
\newcommand{\vctr}[1]{\bm{#1}}
\newcommand{\unitv}[1]{\hat{\vctr{#1}}}
\newcommand{\preup}[1]{\prescript{#1}{}{}}
\newcommand{\rf}[1]{\mathcal{\MakeUppercase{#1}}}
\newcommand{\dcm}[1]{\left[\rf{#1}\right]}
\newcommand{\vrf}[2]{\preup{\rf{#1}}\vctr{#2}}
\newcommand{\mbar}[0]{\;\middle|\;}
\newcommand{\prf}[1]{\preup{\rf{#1}}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

\newcommand{\matcp}[1]{\left[#1 \times\right]}
\newcommand{\rthree}[0]{$\mathbb{R}^3$\:}
\newcommand{\sthree}[0]{$\mathbb{S}^3$\:}
\newcommand{\sothree}[0]{$SO_3$}

@@

\title{Optimal Light Curve Attitude Inversion with Measurement Noise: Two Case Studies}
\author{Liam Robinson}
\author{Carolin Frueh}
\affil{Purdue University, West Lafayette, United States, Email: \texttt{\{robin502, cfrueh\}$@$purdue.edu}}

@@

\FloatBarrier
\section{Inversion Method}

@@

\FloatBarrier
\section{Computing the Light Curve}

@@

\subsection{Computing the Mean Observed Signal}

@@

\subsection{Approximating the Measurement Noise}

The variance of a space object's observed brightness whose mean is determined by Equation \ref{eq:general_bright} is a combination of many independent stochastic processes. These distributions have variances $\sigma^2_\text{sensor}$ due to the sensor's integration and readout effects, $\sigma^2_\text{flat}$ from sensor flat-fielding effects, scintillation noise $\bar{S}_{\text{SO},k} \sigma^2_{Y,k}$ \cite{osborn2015}, Poisson signal shot noise $\lambda_{\text{shot},k}$, and Poisson background noise $\lambda_{\text{back},k}$. The sum of these variances yields the total signal variance in ADU:

@@

\FloatBarrier
\section{Attitude Inversion Results for Synthetic Data}

To demonstrate the effectiveness of this method, we first present results for a simulated light curve where the ground truth is known.

\begin{table}[ht]
  \centering
  \caption{Simulated observation parameters}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  Target COSPAR ID & 1990-054D \\ \hline
  First obs.\ (UTC) & Mar 8, 2025 02:00:00.000 \\ \hline
  Light curve duration & $5$ minutes \\ \hline
  Observations & $100$ \\ \hline
  Integration time & $0.5$ seconds \\ \hline
  \end{tabular}
  \label{tb:case1_in}
\end{table}

\begin{table}[ht]
  \centering
  \caption{Reflection properties of materials used in the synthetic data test case. The authors estimated the white paint parameters qualitatively}
  \vspace*{6pt}
  \begin{tabular}{|l|l|l|l|}
  \hline
  \textbf{Material} & $C_d$ & $C_s$ & $n$ \\ \hline
  Aluminum \cite{fankhauser2023} & $0.34$ & $0.40$ & $8.9$ \\ \hline
  White paint & $0.9$ & $0.1$ & $1$ \\ \hline
  \end{tabular}
  \label{tb:real_matprops}
 \end{table}

\begin{table}[ht]
  \centering
  \caption{True object attitude profile for all test cases}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
 Initial orientation & $\vctr{p}(0) = -\frac{1}{3} \begin{bmatrix} 1 & 1 & 1 \end{bmatrix}^T$ \\ \hline
 Initial $\vctr{\omega}$ ($\text{rad}/\text{s}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \hline
 Inertia tensor ratios & $J_y / J_x = 1, \: J_z / J_x = 0.25$ \\ \hline
 External torque & $\vctr{M}=\begin{bmatrix} 0 & 0 & 0 \end{bmatrix}^T$ \\ \hline
  \end{tabular}
  \label{tb:synth_truth}
\end{table}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figsmall]{rocket_bodies2.png}
  \caption{Measured rocket body model \cite{nasa_models}.}
  \label{fig:meas_model}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figsmall]{rocket_bodies1.png}
  \caption{Simplified rocket body model used for inversion.}
  \label{fig:inv_model}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figsmall]{light_curves_adu_rocket_body.png}
  \caption{Simulated observed flux for the rocket body.}
  \label{fig:obs_adu_synth}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figsmall]{light_curves_mag_rocket_body.png}
  \caption{Simulated observed V-band magnitude for the rocket body.}
  \label{fig:obs_mag_synth}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figsmall]{light_curves_std_rocket_body.png}
  \caption{Simulated observed light curve standard deviation computed with Equation \ref{eq:sigma_total}.}
  \label{fig:obs_std_synth}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figsmall]{sigma_solsRocket Body.png}
  \caption{Converged solution initial orientation MRP vectors.}
  \label{fig:omega_sols1}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figsmall]{omega_sols_Rocket Body.png}
  \caption{Converged solution initial body-frame angular velocities.}
  \label{fig:omega_sols1}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figsmall]{spin_proc_Rocket Body.png}
  \caption{Converged solution body-frame spin and precession rates.}
  \label{fig:omega_sols1}
\end{figure}

\FloatBarrier
\section{Attitude Inversion Results for Real Data}

@@

\begin{table}[ht]
  \centering
  \caption{Delta I rocket body observation parameters}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  Target COSPAR ID & 1984-115C \\ \hline
  First obs.\ (UTC) & Mar 2, 2025 01:53:30.251 \\ \hline
  Light curve duration & $7.5$ minutes \\ \hline
  Observations & $100$ ($92$ extracted) \\ \hline
  Obs.\ frequency & $0.222$ Hz \\ \hline
  Integration time & $3$ seconds \\ \hline
  \end{tabular}
  \label{tb:case1_in}
\end{table}

@@

\begin{table}[ht]
  \centering
  \caption{ECHOSTAR II observation parameters}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  Target COSPAR ID & 1996-055A \\ \hline
  First obs.\ (UTC) & Feb 26, 2025 04:16:12.279 \\ \hline
  Light curve duration & $19.5$ minutes \\ \hline
  Observations & $456$ ($395$ extracted) \\ \hline
  Obs.\ frequency & $0.388$ Hz \\ \hline
  Integration time & $1$ seconds \\ \hline
  \end{tabular}
  \label{tb:case2_in}
\end{table}

@@

\begin{table}[ht]
  \centering
  \caption{ECHOSTAR II inversion assumptions}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Material} & \textbf{Coefficients} \\ \hline
  Aluminum BRDF & $C_d=0.22$, $C_s=0.4$, $n=5$ \\ \hline
  MLI BRDF & $C_d=0.05$, $C_s=0.24$, $n=20$ \\ \hline
  Solar panel BRDF & $C_d=0.05$, $C_s=0.13$, $n=10$ \\ \hline
  $|\omega(0)|_\text{max}$ & $2.32$ $\text{deg} / \text{s}^{-1}$ \\ \hline
  \end{tabular}
  \label{tb:case2_ass}
\end{table}

\FloatBarrier
\section*{Acknowledgments}

@@

\FloatBarrier
\section*{Appendix}

\begin{table}[ht]
  \centering
  \caption{Telescope information common to all test cases}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  CCD sensor & KAF-16803 \\ \hline
  Effective aperture area $A_\circ$ & $0.076$ $m^2$ \\ \hline
  Observer location & $32.900^\circ \textrm{ N}, -105.533^\circ \textrm{ W}$ \\ \hline
  Observer altitude & $2.24$ km above MSL \\ \hline
  Source pixels $n_\text{pix}$ & $100$ \\ \hline
  Read noise $\sigma_\text{read}$ & $9$ $\text{ADU} / \text{pixel}$ \\ \hline
  Dark current rate $\lambda_\text{dark}$ & $0.01$ $\text{ADU} / \text{pixel} / \text{s}$ \\ \hline
  Flat field strength $f_k$ & $0.01$ nondim \\ \hline
  Gain $g$ & $5.1$ $e^- / \text{ADU}$ \\ \hline
  \end{tabular}
  \label{tb:tele_info}
 \end{table}

 \begin{table}[ht]
  \centering
  \caption{BFGS solver configuration for all results cases}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  Maximum function evaluations & $1000$ \\ \hline
  Maximum iterations & $100$ \\ \hline
  Finite difference step size & $1 \cdot 10^{-5}$ \\ \hline
  Gradient $\infty$-norm tolerance & $1 \cdot 10^{-5}$ \\ \hline
  \end{tabular}
  \label{tb:bfgs_info}
 \end{table}

\FloatBarrier
\clearpage
\bibliography{cache/refs.bib}
\end{document}
